draft 1

### 2016-06-19 finalwriteup/draft.rmd
started review of variable selection method

Once

@@

Before determining which variables contribute most to the response variable, return per dollar, outliers were deleted. Specifically, cases with extreme return per dollar values were investigated as special scenarios may have contributed to these values and the model is not intended to predict special scenarios. Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles as intially practised by Tukey, the inventor of the box plot [@Tukey1977]. The outlier projects determined by this method were then reviewed with an employee from the case study company. Based on their domain knowledge and assessment of the outlier cases, cut-off values for the range of return per dollar values were determined. Outliers were projects with return per dollar values greater than 3 or below -2.

@@

Three methods of variable selection were trialled and compared to test whether the different methods highlighted different variables as important. These were Analysis of Variance (ANOVA), conditional inferece forests (cforests), and random forests. It was expected that random forests would highlight variables that had more categories as more important which is a well known bias; it is also a bias that cforests have overcome [@Strobl2007]. ANOVA is based on linear theory which contrasts conditional forest and random forests more free-form structure. Therefore, it is important to compare the contrasting methods' assessment of variable importance.

When running variable importance tests, the dependent variable must be clearly defined. For this problem, the dependent variable can be numeric or categorical: whether a project was profitable (binary classification) or *how* profitable a project was (regression). The regression model was chosen as a variable important for predicting the degree of profitability would also predict profitable vs not profitable.

>>>>

Once variables were engineered, the value each variable could contribute to the predictive model was assessed. As previously stated, variable selection can enhance model accuracy by reducing noise from irrelevant variables and increase simplicity which in turn improves interpretability for the user. To assess variable importance, models predicting 'return per dollar' were built using all available explanatory variables. The models then output which variables contributed significantly or most improved results. Different types of predictive models have different methods of calculating variable importance and since it was not known which model would best predict return per dollar, a few models' variable outputs were reviewed. These included ANOVA linear regression and two ensemble tree methods. Before models assessing variable importance were built, outlier cases were deleted to prevent distortion of valuable explanatory variables.

@@

Case study projects with extreme 'return per dollar' values were likely the result of special scenarios which the model is not intended to predict. For example, many projects appeared wildly profitable but on closer inspection these were all very small jobs. After discussion with the case study company, it was concluded that for those outliers, the small jobs most likely required so little time that busy employees did not bother writing down their hours for that job. Outliers were initially defined as return per dollar values sitting outside 1.5x the interquartile range (IQR) from the upper and lower quartiles. This rule was intially practised by Tukey, the inventor of the box plot [@Tukey1977]. All outlier projects determined by this method were reviewed with an employee from the case study company and based on their domain knowledge and assessment of the outlier cases, reasonable cut-off values for the range of return per dollar values were determined. In summary, projects with 'return per dollar' values greater than 3 or below -2 were removed.

@@

Three methods which output variable importance were trialled and compared to test whether they highlighted different variables as valuable. This was anticipated as each method models data using distinct theoretical foundations. The tested methods were ANOVA linear regression, conditional inference forests (cforests), and random forests. It was expected that random forests would favour variables that had more categories, a well known bias; which is also a bias that cforests have overcome [@Strobl2007]. ANOVA is based on linear theory which contrasts conditional forest and random forest's free-form structure. Therefore, it is benefical to compare the methods' assessments of variable importance.

When running variable importance tests, the dependent variable must be clearly defined. For this problem, the response variable could be continuous or categorical i.e. *how* profitable a project was (continous) or whether a project was profitable or not (binary classification). The continuous response variable was chosen for the variable importance models as the *degree* of profitability also gives the binary classification (profitable or unprofitable).
